### 2018-05-31 Sections/optimising_the_world.tex
Touched up optimisation

To lay the context for the remainder of the section, and illustrate the importance and transformative potential for improved optimisation approaches, we summarise several key everyday applications for optimisation theory that we all rely on, usually unwittingly, executed somewhere in the background up in the cloud, the end user oblivious to the gargantuan calculations being performed behind the scenes merely to update a barely noticeable on-screen logo. There are countless more examples than we have time and space to summarise here.

The examples we present are all examples of \textit{satisfiability problems}\index{Satisfiability problems} -- the problem of trying the simultaneously satisfy a large number of potentially competing constraints in a system.

It is well known that the most complex such satisfiability problems are \textbf{NP}-complete\index{\textbf{NP}-complete} with no known efficient classical solutions (in fact even very trivial constructions of satisfiability problems are already \textbf{NP}-complete, e.g Fig.~\ref{fig:3SAT}).

@@

In such multi-user systems, with so many users, {\sc Greedy}\index{Greedy strategy} strategies, based on optimising users individually are highly sub-optimal, and globally optimised algorithms must be pursued. For example, if there are a thousand drivers competing to get from Town Hall to Redfern, it makes no sense to optimise them individually, since then they will all be identically directed down George St, which will of course quickly saturate and congest. Obviously load-balancing across the various routes benefits everyone collectively \textit{and} individually.

In the case of the discredited, but straightforward {\sc Greedy} strategy, a simple application of Dijkstra's shortest path algorithm (Sec.~\ref{sec:shortest_path}) will yield the optimal route for an individual user, and very quickly since Dijkstra's algorithm has only $O(n^2)$ complexity (i.e a \textbf{P} algorithm).

However, the complexities of a global multi-user optimisation are intuitively evident -- with countless competing interests, finding a set of routes that minimises net transit time yields an enormous space of possibilities and variations to consider, far beyond individual optimisations performed independently. Alas, Dijkstra's tempting algorithm is no longer applicable, and we must employ harder algorithms, many of which are \textbf{NP}-complete. Some of these were introduced in the context of network routing algorithms in Sec.~\ref{sec:route_strats}.

@@

The importance of this problem is obvious. Minimising the resources required to operate public transport effectively could save enormous amounts of money from state budgets. Not to mention, passenger waiting time is of value too. If a million passengers lose just a few minutes of productivity per day to increased commute times, this can amount to billions of dollars a year in lost productivity to the broader economy.

@@

Classical techniques for attacking optimisation problems such as the ones presented above typically come in two flavours:
\begin{itemize}
	\item Heuristics: efficient algorithms that find suboptimal, but satisfactory \textit{approximations} to the solution to a problem. These algorithms are fast and can be efficiently implemented classically, but can have highly variable accuracy in their approximation.
	\item Brute-force: use raw computational power to exhaustively work through all the combinatorics of a problem. This is guaranteed to find the optimal solution, but is typically extremely slow and limited to small instances of the problem.
\end{itemize}

These techniques present us with a tradeoff between the optimality of a solution and the computational resources required to obtain it.

@@

\begin{figure}[!htb]
\includegraphics[width=0.3\textwidth]{3SAT}
\caption{Digital circuit for an instance of the 3-\textsc{SAT} problem, with 4 clauses acting on input variables $\{x_i\}$. Each of the OR gates is input with some combination of 3 of the input bits or their compliments (i.e with a NOT gate). Each of these is referred to as a `clause', of which there are 4 in this example, but could be any number in general. The final AND gate requires that all clauses be simultaneously satisfied in order to yield a final output of `1'. The goal of the problem is to find an input bit-string $x$ that yields an output of `1'. In general, this may require exhaustively searching over the entire space of input states via brute-force, which exhibits time-complexity exponential in the length of the bit-string. This problem is proven to be \textbf{NP}-complete. Note that the similarly-defined 2-\textsc{SAT}\index{2-SAT problem} problem (i.e clauses each contain 2 input bits) is \textbf{P}, making 3-\textsc{SAT} the simplest model to consider in the study of \textbf{NP}-complete problems.} \label{fig:3SAT}	
\end{figure}

@@

Some problems in even harder classes than \textbf{NP}-complete can in some instances be \textit{approximated} using the same approach. The key to solving such problems is to define an oracle\index{Oracle} that attributes a \textit{score} to a given input, rather than a yes/no answer to satisfiability, and answers `yes' or `no' depending on whether that score is above some defined threshold. As an illustrative example, consider the optimisation of, say, a complex traffic network, where the goal is to maximise flow through the network. Then we might define our score to be some flow metric for the network's graph.

@@

\comment{\textbf{BQP} optimisation problems}

@@

While the dream-goals of perfectly optimised resource allocation and perfect economic efficiency are fantasy, we should not underestimate the impact that even modest enhancements in the efficiency of large-scale economic systems could have. While quantum computers will never perfectly solve humanity's immense resource allocation problems, even the modest improvements they offer could be of immense value to the world economy, with enormous implications monetarily, socially, and environmentally. Indeed, o well ptimisation problems are so ubiquitous that pursuing quantum-enhanced optimisation will be of central importance in the quantum era.

>>>>

To lay the context for the remainder of the section, and illustrate the importance and transformative potential for improved optimisation approaches, we summarise several familiar everyday applications for optimisation theory that we all rely on, usually unwittingly, executed somewhere in the background up in the cloud, the end user oblivious to the gargantuan calculations being performed behind the scenes merely to update a barely noticeable on-screen logo. There are countless more examples than we can't possibly have the time and space to comprehensively summarise here.

The examples we present are all examples of \textit{satisfiability problems}\index{Satisfiability problems} -- the problem of trying to simultaneously satisfy a large number of potentially competing constraints in a system, with the best possible outcome.

It is well known that the most complex such satisfiability problems are \textbf{NP}-complete\index{\textbf{NP}-complete} in general, with no known efficient classical solutions (in fact even very trivial constructions of satisfiability problems are already \textbf{NP}-complete, e.g Fig.~\ref{fig:3SAT}).

Here are a mere few notable everyday applications for optimisation theory.

@@

To implement a straightforward {\sc Greedy} strategy, a simple application of Dijkstra's shortest path algorithm (Sec.~\ref{sec:shortest_path}) will yield the optimal route for an individual user, and very quickly since it has only $O(n^2)$ complexity (i.e a \textbf{P} algorithm).

However, in such multi-user systems, with so many users, {\sc Greedy}\index{Greedy strategy} strategies, based on optimising users individually, are highly sub-optimal, and globally optimised algorithms must be pursued. For example, if there are a thousand drivers competing to get from Town Hall to Redfern, it makes no sense to optimise them individually, since then they will all be identically directed down George St, which will of course quickly saturate and congest. Obviously load-balancing across the various routes benefits everyone collectively \textit{and} individually.

The complexities of a global multi-user optimisation are intuitively evident -- with countless competing interests, finding a set of routes that minimises net transit time yields an enormous space of possibilities and variations to consider, far beyond individual optimisations performed independently. Alas, Dijkstra's tempting algorithm is not very useful, and we must employ harder algorithms, many of which are \textbf{NP}-complete. Some of these were introduced in the context of network packet routing algorithms in Sec.~\ref{sec:route_strats}, a conceptually identical problem to road traffic routing.

@@

The importance of this problem is obvious. Minimising the resources required to operate public transport effectively could save enormous amounts of money in state budgets. Not to mention, passenger waiting time is of value too. If a million passengers lose just a few minutes of productivity per day to increased commute times, this can amount to billions of dollars a year in lost productivity to the broader economy.

@@

Classical techniques for attacking optimisation problems such as the ones presented above typically come in two flavours:
\begin{itemize}
	\item Heuristics: efficient algorithms that find suboptimal, but satisfactory \textit{approximations} to the solution to a problem. These algorithms are fast and can be efficiently implemented classically, but can have highly variable accuracy in the solution they provide.
	\item Brute-force: use raw computational power to exhaustively work through all the combinatorics of a problem. This is guaranteed to find the optimal solution, but is typically extremely slow and limited to small instances of the problem.
\end{itemize}

These techniques present us with a tradeoff between the optimality of a solution and the computational resources required to obtain it. Can quantum technology improve this tradeoff?

@@

\begin{figure}[!htb]
\includegraphics[width=0.3\textwidth]{3SAT}
\caption{Digital circuit for an instance of the 3-\textsc{SAT} problem, with 4 clauses acting on input variables $\{x_i\}$. Each of the OR gates is input with some combination of 3 of the input bits or their compliments (i.e with a NOT gate). Each of these is referred to as a `clause', of which there are 4 in this example, but could be any number in general. The final AND gate requires that all clauses be simultaneously satisfied in order to yield a final output of `1'. The goal of the problem is to find an input bit-string $x$ that yields an output of `1'. In general, this may require exhaustively searching over the entire space of input states via brute-force, which exhibits time-complexity exponential in the length of the bit-string. This problem is proven to be \textbf{NP}-complete. Note that the similarly-defined 2-\textsc{SAT}\index{2-SAT problem} problem (i.e clauses each contain 2 input bits) is \textbf{P}, making 3-\textsc{SAT} the simplest model to consider in the study of \textbf{NP}-complete problems. For this reason 3-\textsc{SAT} is often used as the computational model when studying \textbf{NP}-complete problems.} \label{fig:3SAT}	
\end{figure}

@@

Some problems in even harder classes than \textbf{NP}-complete can in some instances be \textit{approximated} using the same approach. The key to solving such problems is to define an oracle\index{Oracle} that attributes a \textit{score}\index{Score} to a given input, rather than a yes/no answer to satisfiability, and answers `yes' or `no' depending on whether that score is above some defined threshold. As an illustrative example, consider the optimisation of, say, a complex traffic network, where the goal is to maximise flow through the network. Then we might define our score to be some flow metric for the network's graph.

@@

\subsection{\textbf{BQP} optimisation problems}\index{\textbf{BQP} optimisation problems}

The optimisation problems discussed until now have been satisfiability problems residing in \textbf{NP}-complete, limited to only quadratic quantum enhancement. Are there any other optimisation problems that natively reside in \textbf{BQP}, and therefore inherently offer exponential quantum speedup? There are, although the examples are far less ubiquitous than the plethora of satisfiability-based optimisation problems.

\comment{To do!}

@@

While the dream-goals of perfectly optimised resource allocation and perfect economic efficiency are fantasy, we should not underestimate the impact that even modest enhancements in the efficiency of large-scale economic systems could have. While quantum computers will never perfectly solve humanity's immense resource allocation problems, even the modest improvements they offer could be of immense value to the world economy, with enormous implications monetarily, socially, and environmentally. Indeed, optimisation problems are so ubiquitous that pursuing quantum-enhanced optimisation will be of central importance in the quantum era.
